---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Load the data.
```{r}
rawdata <- read.table(unz("activity.zip", "activity.csv"), header = TRUE, sep = ",")
```

Process/transform the data into a format suitable for the analysis.
```{r}
# Change the date columns have to be converted to date.
cleandata <- rawdata
cleandata$date <- as.Date(cleandata$date)
```

## What is mean total number of steps taken per day?

In order to calculate the mean total number of daily steps, I created a dataset that consists of date and daily total steps.
```{r}
# Load the dplyr package
library(dplyr)
by_date <- group_by(cleandata, date)
by_date_steps <- summarize(by_date, sum(steps))
by_date_steps <- rename(by_date_steps, totalsteps = `sum(steps)`)
```

```{r histogram, fig.height=4}
hist(by_date_steps$totalsteps, main = paste("Histogram of total number of steps taken each day"), xlab = "Total number of steps taken each day")
```

```{r}
# The mean of the total number of steps taken per day
mean_dailysteps <- mean(by_date_steps$totalsteps, na.rm = TRUE)
mean_dailysteps
# The median of the total number of steps taken per day
median_dailysteps <- median(by_date_steps$totalsteps, na.rm = TRUE)
median_dailysteps
```

## What is the average daily activity pattern?

```{r}
by_interval <- group_by(cleandata, interval)
by_interval_steps <- summarize(by_interval, mean(steps, na.rm = TRUE))
by_interval_steps <- rename(by_interval_steps, averagesteps = `mean(steps, na.rm = TRUE)`)
plot(by_interval_steps, type = "l")
```

```{r}
max_steps_interval <- filter(by_interval_steps, averagesteps == max(averagesteps))
max_steps_interval
```

The 5-minute interval of `r max_steps_interval$interval` contains the maximum number of steps.

## Imputing missing values
The total number of the missing values in the dataset is calculated as below.
```{r}
sum(is.na(cleandata$steps))
```

Taking into consideration that the dataset contains a large number of observations, I decided to use the mean (instead of median) for the interval to fill the missing values in the dataset.

```{r}
# Extract data with NAs and fill them with the mean for the interval
nadata <- cleandata[is.na(cleandata),]
nadata$steps <- by_interval_steps$averagesteps
# Merge the dataset
nonnadata <- na.omit(cleandata)
filleddata <- rbind(nadata, nonnadata)
```

I made a histogram of the total number of steps taken each day and calculated the mean and median total number of steps taken per day as follows.

```{r}
by_date2 <- group_by(filleddata, date)
by_date_steps2 <- summarize(by_date2, sum(steps))
by_date_steps2 <- rename(by_date_steps2, totalsteps = `sum(steps)`)

hist(by_date_steps2$totalsteps, main = paste("Histogram of total number of steps taken each day (with NAs filled in)"), xlab = "Total number of steps taken each day")
```

```{r}
# The mean of the total number of steps taken per day
mean_dailysteps2 <- mean(by_date_steps2$totalsteps)
mean_dailysteps2
# The median of the total number of steps taken per day
median_dailysteps2 <- median(by_date_steps2$totalsteps)
median_dailysteps2
```

The mean value remain the same as the first part of the assignment, but the median value became the same value as the mean value in the second analysis. Since the NAs are filled with the mean values of the interval, it wouldn't affect the mean value.

## Are there differences in activity patterns between weekdays and weekends?
